Tailoring Differentially Private Bayesian Inference to Distance Between Distributions

### 2018-10-21 Arxiv/main.tex
update arvix paper

\section{Algorithm Setting up}
\label{sec_setup}
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and a sequence of $n$ observations $\dataobs\in\{0,1\}^n$, we define the follwing set:

@@

\subsection{Baseline Approaches}

\subsubsection{Exponential Mechanism}

Exponential mechanism $\expmech{}{x}{u}{\candidateset}$ samples a element from the candidate set $\candidateset = \{r_1, r_2, \cdots r_n\}$ with probability proportional to $exp(\frac{\epsilon u(x,r)}{2 GS})$:
\begin{equation*}
\underset{z \thicksim \expmech{}{x}{u}{\candidateset} }{Pr}[z=r] = \frac
{exp(\frac{\epsilon u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\ exp(\frac{\epsilon u(x,r')}{2 GS})},
\end{equation*}
where $u(x,r)$ is the Hellinger scoring function over candidates, $\hellinger(\bysinfer(\dataobs), r)$, and $GS(x)$ is the global sensitivity calculated by:
\begin{equation*}
GS = 
\max_{\{|\dataobs,\dataobs'| \leq 1;\dataobs,\dataobs'\in \mathcal{X}^n\}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}

@@

\subsection{Our approach: smoothed Hellinger distance based exponential mechanism}
\label{sec_smoo}

@@

In this section, we evaluate the accuracy of the mechanisms defined in
Section (\ref{sec_smoo}) w.r.t. four variables, including data size, dimensions,
data variance, prior distribution, and some combinations thereof.
Every plot is an average over 1000 runs. In all the experiments we set
$\epsilon = 0.8$, and $\delta = 10^{-8}$.

\noindent In the following some of the plots show
mean error as a function of the datasize while one
is a whiskers-plot where the y-axis shows the average
accuracy (or equivalently, the error) of the mechanisms, and the x-axis, instead shows
different balanced priors used. The boxes extend from the lower to the upper quartile values
of the data, with a line at the median. A notch on the box around the
median is also drawn to give a rough guide to the significance of
difference of medians; The whiskers extend from the box to show the
range of the data. A blue box in the plots represents our newly
designed exponential mechanism's behavior-- where the sensitivity is calibrated
w.r.t Hellinger distance-- while the yellow box next to
it represents the performance of a variation of the basic Laplace
mechanism presented in Section (\ref{sec:base}) with the same
settings: that is $\epsilon, \delta$, data, prior. The variation
considered performs a postprocessing on the released parameters so
that they are consistent. For instance when the sum of the noised
parameters is greater than $n$ we will truncate them so that they sum
up to $n$.

@@

\begin{figure}[H]
\begin{center}
\centering
  \subfigure[Data set size from $300$ to $800$ ]{
    \includegraphics[width=0.45\textwidth]{sampling_2d.eps}
    \label{subsubsec_vs_datasize1a}
  }
\caption{Increasing data size with fixed prior $\betad(1,1)$. Unbalanced datasets of mean $(0.1,0.9)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
  \subfigure[Data set size from $14000$ to $20000$]{
     \includegraphics[width=0.45\textwidth]{sampling_3d.eps}   
  \label{subsubsec_vs_datasize1bdir}
  } 
\caption{Increasing data size with fixed $\dirichlet(1,1,1)$ prior distribution, Unbalanced datasets of mean $(0.2,0.3, 0.5)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_dir}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
  \subfigure[Data set size from $14000$ to $20000$]{
     \includegraphics[width=0.45\textwidth]{sampling_4d.eps}   
  \label{subsubsec_vs_datasize1bdir}
  } 
\caption{Increasing data size with fixed $\dirichlet(1,1,1)$ prior distribution, Unbalanced datasets of mean $(0.2,0.3, 0.5)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_dir}
\end{center}
\end{figure}

In Figures \ref{fig_vs_datasize} and \ref{fig_vs_datasize_dir} we consider \emph{unbalanced} datasets
of observations. This means that in the Beta-Binomial setting (Figure \ref{subsubsec_vs_datasize1a},
and \ref{subsubsec_vs_datasize1b}) the datasets will consist of 10\% 1s and the rest 0s, while for the
Dirichelet-Multinomial (Figure  \ref{subsubsec_vs_datasize1adir} and \ref{subsubsec_vs_datasize1bdir})
the data will be split in the $k=3$ bins with perecentages of: 20\%, 30\% and 50\%.
The results show that when the data size
increases, the average errors of
$\hexpmech$, $\hexpmechd$, and Laplace decrease. For small datasets,
i.e with size less 650 in the case of Beta-Binomial systems,
the Laplace mechanisms outperforms the exponential mechanisms,
but for bigger data sets, that is, bigger than 650, or as in Figure \ref{subsubsec_vs_datasize1b} where
we considered data sets of the order of 15 thousands elements,
the exponential mechanisms outperforms the Laplace mechanism.
Similar experimental tendencies were obtained for the Dirichlet-Multinomial system ( (Figure  \ref{subsubsec_vs_datasize1adir} and \ref{subsubsec_vs_datasize1bdir})),
but we were not able to perform experiments with higher dimensions (4 or more) or, big datasets, e.g. $10^4$ elements, due to
too high run time of the algorithm.

\paragraph{Fixed dataset varying balanced priors}
\label{subsubsec_vs_prior}
In Figure \ref{fig_vs_prior}, we fix the dataset to be $\langle 5,5,5\rangle$.
We also considered balanced priors with increasing values in their dimensions.
The plot shows that in the beginning the Laplace mechanism performs better but
it is outperformed after a while.
\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{sampling_prior.eps}
\caption{Observed data set is: $(5,5,5)$, varying balanced priors}
\label{fig_vs_prior}
\end{figure}

\subsection{Experiment Evaluations on Privacy}
\label{subsec_experiment_privacy}
In order to see our privacy behavior, we study the accurate epsilon under concrete cases in this section. The $(\epsilon, \delta)$ - differential privacy we proved in Sec. \ref{sec_smoo} is just an upper bound, we concrete $\epsilon$ should be smaller than upper bound in our exponential mechanism. We calculate the concrete privacy value in following ways wrt. the data size, and obtain plots in Fig. \ref{fig_privacy}.

\begin{figure*}
\begin{center}
\centering
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20.eps}}
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20(2).eps}} 
\caption{Concrete privacy calculation under settings that: prior distribution:$[1,1]$, $\epsilon = 0.8$, $\delta = 0.0005$ and observed data are uniformly distributed}
\label{fig_privacy}
\end{center}
\end{figure*}

>>>>

\section{Mechanisms Introduction}
\label{sec_mechs}
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and a sequence of $n$ observations $\dataobs\in\{0,1\}^n$, we define the follwing set:

@@

\subsection{Baseline Approaches}
\label{subsec_baselines}

\subsubsection{Exponential Mechanism}
Exponential mechanism $\expmech{}{x}{u}{\candidateset}$ samples a element from the candidate set $\candidateset = \{r_1, r_2, \cdots r_n\}$ with probability proportional to $exp(\frac{\epsilon u(x,r)}{2 GS})$:
\begin{equation*}
\underset{z \thicksim \expmech{}{x}{u}{\candidateset} }{Pr}[z=r] = \frac
{exp(\frac{\epsilon u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\ exp(\frac{\epsilon u(x,r')}{2 GS})},
\end{equation*}
where $u(x,r)$ is the Hellinger scoring function over candidates, $\hellinger(\bysinfer(\dataobs), r)$, and $GS(x)$ is the global sensitivity calculated by:
\begin{equation*}
GS = 
\max_{\{|\dataobs,\dataobs'| \leq 1;\dataobs,\dataobs'\in \mathcal{X}^n\}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}

@@

\subsection{Our approach: smoothed Hellinger distance based exponential mechanism}
\label{subsec_hexpmech}

@@

In this section, we evaluate the accuracy of the mechanisms defined in
Section (\ref{sec_mechs}) w.r.t. four variables, including data size, dimensions,
data variance, prior distribution, and some combinations thereof.
Every plot is an average over 1000 runs. In all the experiments we set
$\epsilon = 1.0$, and $\delta = 10^{-8}$.

\noindent In the following some of the plots show
mean error as a function of the datasize while one
is a whiskers-plot where the y-axis shows the average
accuracy (or equivalently, the error) of the mechanisms, and the x-axis, instead shows
different balanced priors used. The boxes extend from the lower to the upper quartile values
of the data, with a line at the median. A notch on the box around the
median is also drawn to give a rough guide to the significance of
difference of medians; The whiskers extend from the box to show the
range of the data. A blue box in the plots represents our newly
designed exponential mechanism's behavior-- where the sensitivity is calibrated
w.r.t Hellinger distance-- while the yellow box next to
it represents the performance of a variation of the basic Laplace
mechanism presented in Section (\ref{subsec_baselines}) with the same
settings: that is $\epsilon, \delta$, data, prior. The variation
considered performs a postprocessing on the released parameters so
that they are consistent. For instance when the sum of the noised
parameters is greater than $n$ we will truncate them so that they sum
up to $n$.

@@

\begin{figure}[H]
\begin{center}
\centering
    \includegraphics[width=0.48\textwidth]{sampling_2d.eps}
\caption{Increasing data size with prior $\betad(1,1)$, balanced datasets and parameters $\epsilon = 1.0$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_2d}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
     \includegraphics[width=0.48\textwidth]{sampling_3d.eps}   
\caption{Increasing data size with $\dirichlet(1,1,1)$ prior distribution, balanced datasets and parameters $\epsilon = 1.0$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_3d}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
     \includegraphics[width=0.48\textwidth]{sampling_4d.eps}    
\caption{Increasing data size with $\dirichlet(1,1,1,1)$ prior distribution, Unbalanced datasets and parameters $\epsilon = 1.0$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_4d}
\end{center}
\end{figure}

In Figures \ref{fig_vs_datasize_2d}, \ref{fig_vs_datasize_3d} and \ref{fig_vs_datasize_4d} we consider \emph{balanced} data sets
of observations. This means that in the Beta-Binomial setting the datasets will consist of 50\% 1s and the rest 0s, while for the
Dirichelet-Multinomial (Figure  \ref{fig_vs_datasize_3d} and \ref{fig_vs_datasize_4d})
the data will be split in the $k=3$ bins with perecentages of: 33\%, 33\% and 34\% in 3 dimensionality and 25\%, 25\%, 25\% and 25\% in 4 dimensionality.
The results show that when the data size increases, the average errors of
$\hexpmech$, Laplace mechanism and decrease. For small datasets,
i.e with size less $300$ in the case of Beta-Binomial models,
both the baseline Laplace mechanisms and improved Laplace mechanism outperform $\hexpmech$.
But for bigger data sets, that is, bigger than $300$, or as in Figure \ref{fig_vs_datasize_2d} where
we considered data sets of the order of 15 thousands elements,
the $\hexpmech$ outperforms the baseline Laplace mechanism, and asymptotically approaches the improved baseline Laplace mechanism.
Similar experimental tendencies were obtained for the Dirichlet-Multinomial system (Figure \ref{fig_vs_datasize_3d} and \ref{fig_vs_datasize_4d}).

\paragraph{Fixed dataset varying balanced priors}
\label{subsubsec_vs_prior}
In Figure \ref{fig_vs_prior}, we fix the data set to be $(50,50)$, and the parameters the same as before: $\epsilon = 1.0$ and $\delta = 10^{-8}$ . We studied the accuracy under different priors, where the priors considered  are also balanced.
The plot shows that in the beginning the Laplace mechanism performs better but
it is outperformed after a while.
\begin{figure}[H]
\centering
\includegraphics[width=0.48\textwidth]{sampling_prior.eps}
\caption{Observed data set is: $(50,50)$, varying balanced priors}
\label{fig_vs_prior}
\end{figure}

\subsection{Experiment Evaluations on Privacy Loss}
\label{subsec_experiment_privacy}
In order to see our privacy behavior, we study the accurate epsilon under concrete cases in this section. The $(\epsilon, \delta)$ - differential privacy we proved in Sec. \ref{subsec_hexpmech} is just an upper bound, we concrete $\epsilon$ should be smaller than upper bound in our exponential mechanism. We calculate the concrete privacy value in following ways wrt. the data size, and obtain plots in Fig. \ref{fig_privacy}.

\begin{figure*}
\begin{center}
\centering
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20.eps}}
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20(2).eps}} 
\caption{Concrete privacy calculation under settings that: prior distribution:$[1,1]$, $\epsilon = 1.0$, $\delta = 0.0005$ and observed data are uniformly distributed}
\label{fig_privacy}
\end{center}
\end{figure*}
